---
title: Peer Assesment 1
author: "Gita"
date: "9 February 2015"
output: html_document
---

###Loading and preprocessing the data

```{r echo=TRUE}

mainData <- read.csv("activity.csv", sep=",")
head(mainData)

hist(mainData$steps)
```

###What is mean total number of steps taken per day?

1. Total number of steps calculated for each day and saved in the variable
```{r echo=TRUE}
stepsPerDay <- tapply(mainData$steps, mainData$date, sum, na.rm=TRUE)
head(stepsPerDay)
```



2. Histogram of total number of steps taken each day . 

```{r echo=FALSE}
library(ggplot2)
library(scales)
library(Hmisc)
```

```{r}
qplot(stepsPerDay, xlab="Total steps for day", ylab="Frequency of steps",  binwidth=1000)
```

3. Calculating here the mean and median of total number of steps taken each day.

```{r}
meanByDay <- mean(stepsPerDay)
medianByDay <- median(stepsPerDay)
```

Mean of total number of steps taken each day is ***`r meanByDay`***.  
Mean of total number of steps taken each day is ***`r medianByDay`***.

###What is the average daily activity pattern?

```{r}
averageStepsPerTimeBlock <- aggregate(x=list(meanSteps=mainData$steps), by=list(interval=mainData$interval), FUN=mean, na.rm=TRUE)

head(averageStepsPerTimeBlock)
```

1. Make a time series plot

```{r}

qplot(interval, meanSteps, data=averageStepsPerTimeBlock, geom=c("line"), xlab="Average no of steps taken", ylab="5-min interval")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
mostSteps <- which.max(averageStepsPerTimeBlock$meanSteps)
timeMostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageStepsPerTimeBlock[mostSteps,'interval'])
```
Most Steps at: `r timeMostSteps`  

##Imputing missing values

1. Calculate and report the total number of missing values 

```{r}
missingValues <- length(which(is.na(mainData$steps)))
```

Total number of missing value is ***`r missingValues`***.

2. Filling in all of the missing values in the dataset. 
3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

   Hmisc package impute function is used here.

```{r}
dataImputed <- mainData
dataImputed$steps <- impute(mainData$steps, fun=mean)
```

4. Make a histogram of the total number of steps taken.

```{r}
stepsPerDayImputed <- tapply(dataImputed$steps, dataImputed$date, sum)
qplot(stepsPerDayImputed, xlab='Total steps per day (Imputed)', ylab='Frequency using', binwidth=100)
```

 Calculate and report the mean and median
 
```{r}
stepsPerDayMeanImputed <- mean(stepsPerDayImputed)
stepsPerDayMedianImputed <- median(stepsPerDayImputed)

```

Mean (Imputed): `r stepsPerDayMeanImputed`
Median (Imputed): `r stepsPerDayMedianImputed`

###Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable weekday and weekend

```{r}
dataImputed$dateType <-  ifelse(as.POSIXlt(dataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```
2. Make a time series plot 

```{r}
avgdataImputed <- aggregate(steps ~ interval + dateType, data=dataImputed, mean)
ggplot(avgdataImputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```
